---
title: "ODE with Jacobian"
output: html_notebook
---

## Solving with C 
This is the C module. It is saved as `mymod.c`.
We will compile it below:

```
      /* file mymod.c */
      #include <R.h>
      static double parms[3];
      #define k1 parms[0]
      #define k2 parms[1]
      #define k3 parms[2]
      
      /* initializer  */
      void initmod(void (* odeparms)(int *, double *))
      {
          int N=3;
          odeparms(&N, parms);
      }
      
      /* Derivatives and 1 output variable */
      void derivs (int *neq, double *t, double *y, double *ydot,
                   double *yout, int *ip)
      {
          if (ip[0] <1) error("nout should be at least 1");
          ydot[0] = -k1*y[0] + k2*y[1]*y[2];
          ydot[2] = k3 * y[1]*y[1];
          ydot[1] = -ydot[0]-ydot[2];
      
          yout[0] = y[0]+y[1]+y[2];
      }
      
      /* The Jacobian matrix */
      void jac(int *neq, double *t, double *y, int *ml, int *mu,
                 double *pd, int *nrowpd, double *yout, int *ip)
      {
        pd[0]               = -k1;
        pd[1]               = k1;
        pd[2]               = 0.0;
        pd[(*nrowpd)]       = k2*y[2];
        pd[(*nrowpd) + 1]   = -k2*y[2] - 2*k3*y[1];
        pd[(*nrowpd) + 2]   = 2*k3*y[1];
        pd[(*nrowpd)*2]     = k2*y[1];
        pd[2*(*nrowpd) + 1] = -k2 * y[1];
        pd[2*(*nrowpd) + 2] = 0.0;
      }
      /* END file mymod.c */

```

### Compiling the `C` module
```{r}
# Unload if the the module is loaded
if(is.loaded("mymod")) dyn.unload(paste("mymod", .Platform$dynlib.ext, sep = ""))
# delete any existing files first
file.remove("mymod.o")
file.remove("mymod.dll")
file.remove("mymod.so")

# now, compile
system("R CMD SHLIB mymod.c")
```


### Build the DE model and Jacobian in pure R
```{r}
# model in R
model <- function(t, Y, parameters) {
  with (as.list(parameters),{

    dy1 = -k1*Y[1] + k2*Y[2]*Y[3]
    dy3 = k3*Y[2]*Y[2]
    dy2 = -dy1 - dy3

    list(c(dy1, dy2, dy3))
  })
}

# the Jacobian
jac <- function (t, Y, parameters) {
  with (as.list(parameters),{

    PD[1,1] <- -k1
    PD[1,2] <- k2*Y[3]
    PD[1,3] <- k2*Y[2]
    PD[2,1] <- k1
    PD[2,3] <- -PD[1,3]
    PD[3,2] <- k3*Y[2]
    PD[2,2] <- -PD[1,2] - PD[3,2]

    return(PD)
  })
}

```


### Solve the DifEq with R
```{r}
library(deSolve)

# input parameters
parms <- c(k1 = 0.04, k2 = 1e4, k3 = 3e7)
Y     <- c(1.0, 0.0, 0.0)
times <- c(0, 0.4*10^(0:11))

# PD    <- matrix(nrow = 300, ncol = 300, data = 0)

out   <- ode(Y, times, model, parms = parms, jacfunc = jac)

plot(out)
```

### Load the C module and compare against R
```{r}
# system("R CMD SHLIB mymod.c")

dyn.load(paste("mymod", .Platform$dynlib.ext, sep = ""))

## benchmarks
lim <- 10

# calling the R code
system.time(
  for(i in 1:lim) out   <- ode(Y, times, model, parms = parms)
)

system.time(
  for(i in 1:lim) out   <- ode(Y, times, model, parms = parms, jacfunc = jac)
)

# calling the C module
system.time(
  for(i in 1:lim)
    out <- ode(Y, times, func = "derivs", parms = parms,
               dllname = "mymod",
               initfunc = "initmod", nout = 1, outnames = "Sum")
)

system.time(
  for(i in 1:lim)
    out <- ode(Y, times, func = "derivs", parms = parms,
               jacfunc = "jac", dllname = "mymod",
               initfunc = "initmod", nout = 1, outnames = "Sum")
)


# calling the Fortran module

dyn.load(paste("mymodf", .Platform$dynlib.ext, sep = ""))

system.time(
  for(i in 1:lim)
    out <- ode(Y, times, func = "derivs", parms = parms,
               dllname = "mymodf",
               initfunc = "initmod", nout = 1, outnames = "Sum")
)

system.time(
  for(i in 1:lim)
    out <- ode(Y, times, func = "derivs", parms = parms,
               jacfunc = "jac", dllname = "mymodf",
               initfunc = "initmod", nout = 1, outnames = "Sum")
)
```

## Benchmarking

### Function for benchmarking plots for all functions
```{r}
library(ggplot2)

# build function to plot timings
# devtools::install_github("WinVector/WVPlots")
# library('WVPlots')
plotTimings <- function(timings) {
  timings$expr <- reorder(timings$expr, -timings$time, FUN=max)
  # ggplot(data = timings, aes(x=stress,y=time, color=expr)) +
  #   geom_point(alpha=0.8) + geom_smooth(alpha=0.8) 
  
  nmax <- max(timings$stress)
  tsub <- timings[timings$stress==nmax,]
  tsub$expr <- reorder(tsub$expr,tsub$time,FUN=median)
  list(
    ggplot(data=timings,aes(x=stress,y=time,color=expr)) +
      geom_point(alpha=0.8) + geom_smooth(alpha=0.8),
    ggplot(data=timings,aes(x=stress,y=time,color=expr)) +
      geom_point(alpha=0.8) + geom_smooth(alpha=0.8) +
      scale_y_log10(),
    WVPlots::ScatterBoxPlot(tsub,'expr','time',
                            title = paste('stress = ', nmax)) +

      coord_flip()
  )
}
```


### Functions to test
```{r}
library(microbenchmark)
library(deSolve)

# input parameters
parms <- c(k1 = 0.04, k2 = 1e4, k3 = 3e7)
Y     <- c(1.0, 0.0, 0.0)
times <- c(0, 0.4*10^(0:11))

ode.r.1 <- function(n) {
  for (i in 1:n) ode(Y, times, model, parms = parms)
}

ode.r.2 <- function(n) {
  for (i in 1:n) ode(Y, times, model, parms = parms, jacfunc = jac)
}


ode.c.1 <- function(n) {
  for (i in 1:n)  ode(Y, times, func = "derivs", parms = parms,
               dllname = "mymod",
               initfunc = "initmod", nout = 1, outnames = "Sum")
}

ode.c.2 <- function(n) {
    for (i in 1:n)  ode(Y, times, func = "derivs", parms = parms,
               jacfunc = "jac", dllname = "mymod",
               initfunc = "initmod", nout = 1, outnames = "Sum")
}  

ode.f.1 <- function(n) {
  for (i in 1:n)  ode(Y, times, func = "derivs", parms = parms,
               dllname = "mymodf",
               initfunc = "initmod", nout = 1, outnames = "Sum")
}

ode.f.2 <- function(n) {
    for (i in 1:n)  ode(Y, times, func = "derivs", parms = parms,
               jacfunc = "jac", dllname = "mymodf",
               initfunc = "initmod", nout = 1, outnames = "Sum")
}
```


### quick test with sapply
```{r}
timeSeq <- seq.int(10, 100, 20)
sapply(timeSeq, ode.r.1)
```

### set the test stress
```{r}
kt <- 1
timeSeq <- seq.int(10, 100, 10)             # short sequence
# timeSeq <- rep(10, 11)
timingsPrev <- NULL
```

### test the R function #1
```{r}
# ODE R # 1
# this list will save the dataframes created during the test
timings <- vector("list", length(timeSeq))    # create a vector of lists

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.r.1(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}

timings <- data.table::rbindlist(timings) # convert DF to DT
print(plotTimings(timings))
timingsPrev <- timings                           # save previous timings
```

### Test the R function #2
```{r}
timings <- vector("list", length(timeSeq))    # create a vector of lists

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.r.2(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}

timings <- data.table::rbindlist(timings) # convert DF to DT
print(plotTimings(timings))
```

```{r}
timings <- rbind(timings, timingsPrev)
print(plotTimings(timings))
timingsPrev <- timings                           # accumulate previous timings
```

```{r}
# ODE C #1
timings <- vector("list", length(timeSeq))    # create a vector of lists

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.c.1(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}

timings <- data.table::rbindlist(timings) # convert DF to DT
print(plotTimings(timings))
```

```{r}
timings <- rbind(timings, timingsPrev)
print(plotTimings(timings))
timingsPrev <- timings                           # accumulate previous timings
```

```{r}
# ODE C #2
timings <-  vector("list", length(timeSeq))

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.c.2(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}

timings <- data.table::rbindlist(timings) # convert DF to DT
print(plotTimings(timings))
```


```{r}
timings <- rbind(timings, timingsPrev)
print(plotTimings(timings))
timingsPrev <- timings                           # accumulate previous timings
```


```{r}
# ODE F #1
timings <- vector("list", length(timeSeq))    # create a vector of lists

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.f.1(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}

timings <- data.table::rbindlist(timings) # convert DF to DT
print(plotTimings(timings))
```


```{r}
timings <- rbind(timings, timingsPrev)
print(plotTimings(timings))
timingsPrev <- timings                           # accumulate previous timings
```

```{r}
# ODE F #2
timings <-  vector("list", length(timeSeq))

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.f.2(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}

timings <- data.table::rbindlist(timings) # convert DF to DT
print(plotTimings(timings))
```


```{r}
timings <- rbind(timings, timingsPrev)
print(plotTimings(timings))
timingsPrev <- timings                           # accumulate previous timings
```


## Show only final results
```{r}
kt <- 1
timeSeq <- seq.int(10, 100, 20)             # short sequence
timingsPrev <- NULL

# ODE R # 1
# this list will save the dataframes created during the test
timings <- vector("list", length(timeSeq))    # create a vector of lists

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.r.1(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}
timings <- data.table::rbindlist(timings) # convert DF to DT

timings <- rbind(timings, timingsPrev)
timingsPrev <- timings                           # accumulate previous timings


# ODE R #2
timings <- vector("list", length(timeSeq))    # create a vector of lists

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.r.2(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}
timings <- data.table::rbindlist(timings) # convert DF to DT

timings <- rbind(timings, timingsPrev)
timingsPrev <- timings                           # accumulate previous timings

# ODE C #1
timings <- vector("list", length(timeSeq))    # create a vector of lists

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.c.1(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}
timings <- data.table::rbindlist(timings) # convert DF to DT

timings <- rbind(timings, timingsPrev)
timingsPrev <- timings                           # accumulate previous timings

# ODE C #2
timings <-  vector("list", length(timeSeq))

for(i in seq_len(length(timeSeq))) {          # iterate throgh time sequence
  stress <- timeSeq[[i]]                      # get stress for the current DF
  ti <- microbenchmark(ode.c.2(stress), times = 10)
  ti$stress <- stress                         # document stress
  ti <- data.frame(ti, stringsAsFactors = FALSE) # no factors
  timings[[i]] <- ti                          # save DF in the list
}

timings <- data.table::rbindlist(timings) # convert DF to DT

timings <- rbind(timings, timingsPrev)
print(plotTimings(timings))

```


### Compiling the `C` module
```{r}
# Unload if the the module is loaded
if(is.loaded("mymodf")) dyn.unload(paste("mymodf", .Platform$dynlib.ext, sep = ""))
# delete any existing files first
file.remove("mymodf.o")
file.remove("mymodf.dll")
file.remove("mymodf.so")

# now, compile
system("R CMD SHLIB mymodf.f")
```


```{r}

```

